Supplementary Information: Quantifying the pathways to life using assembly spaces

### 2021-05-14 papers/assembly/I/main.tex
Expand text and examples through pathway connectedness

As a more elaborate example, one which we will repeatedly refer to as we expand on the formalism, we imagine a simple ligation type process where in the characters $A$ and $B$ can be combined into strings and $C$ acts something like a \red{``catalyst''} making some transitions possible if and only if $C$ exists.
Consider the Petri $P$ net described by $S(P) = \{ A, B, C, AA, AB, AAAB \}$ and $T(P) = \{ \alpha, \beta, \gamma, \delta \}$ with source and target maps:

@@

@@

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition, fill=green!20] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

@@

\begin{lemma}\label{lem:image-of-pathways}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $x, y \in S(P)$ be pathway connected in $P$.
    Then $f(x)$ and $f(y)$ are pathway connected in $f(P)$.
\end{lemma}
\begin{proof}
    Let $\gamma = t_n \ldots t_1$ be a pathway from $x$ to $y$.
    Given that $\sigma_P(t_i) \ne 0$ and $\tau_P(t_i) \ne 0$, we know that $\sigma_{f(P)} \ne 0$ and $\tau_{f(P)} \ne 0$.
    Since $\sigma_P(t_1) = x + m$, $\sigma_{f(P)}(f(x)) = f(x) + m'$ for some $m' \in \N[S(f(P))]$; similarly, $\tau_{f(P)}(f(y)) = f(y) + n'$ for some $n' \in \N[S(f(P))]$.

>>>>

As a more elaborate example, one which we will repeatedly refer to as we expand on the formalism as it exhibits number of interesting features, we imagine a simple ligation type process where in the characters $A$ and $B$ can be combined into strings and $C$ acts something like a \red{``catalyst''} making some transitions possible if and only if $C$ exists.
Consider the Petri $P$ net described by $S(P) = \{ A, B, C, AA, AB, AAAB \}$ and $T(P) = \{ \alpha, \beta, \gamma, \delta \}$ with source and target maps:

@@

Up to this point, all of the defintions have been fairly standard, at least in terms of concepts.
We now start constructing apparently new ideas that will allow us to formalize the concepts of an assembly space and the assembly index.
The first of these is the idea of a pathway.
Suppose you begin with a given object $x$ and wish to fleshout all of the ways in which a second object $y$ might be constructed.
The simplest first step is to construct a sequence of transitions, the first of which takes $x$ as part of its input, the last yields $y$ as part of it its output, and every transition in between depends at least partially on the output of the previous.
Such a sequence of transitions is what we refer to as a \textit{pathway}.

@@

It is clear that a pathway is a directed concept in general.
A pathway from $x$ to $y$ is not necessarily a pathway from $y$ to $x$, and the existence of a pathway from $x$ to $y$ does not imply a pathway from $y$ to $x$ exists.
If there is a pathway from $x$ to $y$ we say \textit{$x$ pathway connected to $y$}.
In general, if $x$ is pathway connected to $y$ there may be multiple pathways, and it will be useful to refer to that set of pathways.

@@

In our ligation example, you might suppose that you begin with $A$ and would like to ultimately yield $AAAB$.
There are in fact two pathways which accomplish this, $\gamma\alpha$ and $\gamma\beta$, which we read as \textit{first fire $\alpha$ ($\beta$, respectively) and then fire $\gamma$}:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition, fill=green!20] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
                \node [scale=3] at (5, -3.5) {$\gamma\alpha$};

                \node [style=site] (A2) at (14, 0.8) {$A$};
                \node [style=site] (B2) at (14, -0.8) {$B$};
                \node [style=site] (C2) at (20, -1.6) {$C$};
                \node [style=site] (AA2) at (18, 0.8) {$AA$};
                \node [style=site] (AB2) at (18, -0.8) {$AB$};
                \node [style=site] (AAAB2) at (22, 0) {$AAAB$};
                \node [style=transition] (tau12) at (16, 0.8) {$\alpha$};
                \node [style=transition, fill=green!20] (tau22) at (16, -0.8) {$\beta$};
                \node [style=transition, fill=green!20] (tau32) at (20, 0) {$\gamma$};
                \node [style=transition] (tau42) at (24, 0) {$\delta$};
                \node [scale=3] at (19, -3.5) {$\gamma\beta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);

                \draw [style=arrow, bend right=15] (A2) to (tau12);
                \draw [style=arrow, bend left=15] (A2) to (tau12);
                \draw [style=arrow, bend right=15] (A2) to (tau22);
                \draw [style=arrow] (tau12) to (AA2);
                \draw [style=arrow] (B2) to (tau22);
                \draw [style=arrow] (tau22) to (AB2);
                \draw [style=arrow] (AA2) to (tau32);
                \draw [style=arrow] (AB2) to (tau32);
                \draw [style=arrow, bend left=30] (C2) to (tau32);
                \draw [style=arrow] (tau32) to (AAAB2);
                \draw [style=arrow, bend left=30] (tau32) to (C2);
                \draw [style=arrow] (AAAB2) to (tau42);
                \draw [style=arrow, bend right=30] (tau42) to (A2);
                \draw [style=arrow, bend right=30] (tau42) to (AA2);
                \draw [style=arrow, bend left=40] (tau42) to (B2);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
where the transitions involved in the pathway are colored green.
The set of pathways from $A$ to $AAAB$ is then $\Gamma_{A}^{AAAB} = \{ \gamma\alpha, \gamma\beta \}$.
In the case where there are loops in the Petri net, some or all of the sets of pathways will be countably infinite.
While this seems undesirable, it does not present much of a problem for the problems under consideration in this work.

As with subnets, pathways interact well with Petri net morphisms.
This brings us to our first lemma:
\begin{lemma}\label{lem:image-of-pathways}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $x, y \in S(P)$ be pathway connected in $P$.
    Then $f(x)$ and $f(y)$ are pathway connected in $f(P)$.
\end{lemma}
\begin{proof}
    Let $\gamma = t_n \ldots t_1$ be a pathway from $x$ to $y$.
    Given that $\sigma_P(t_i) \ne 0$ and $\tau_P(t_i) \ne 0$, we know that $\sigma_{f(P)} \ne 0$ and $\tau_{f(P)} \ne 0$.
    Since $\sigma_P(t_1) = x + m$, $\sigma_{f(P)}(f(x)) = f(x) + m'$ for some $m' \in \N[S(f(P))]$; similarly, $\tau_{f(P)}(f(y)) = f(y) + n'$ for some $n' \in \N[S(f(P))]$.
